---
title: "IRTM_CFA_Sims"
author: "Margaret J. Foster"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Document to compare IRT- M and CFA

We designed simulations to compare IRT-M with PCA and unconstrained IRT models.
In this section we extend the comparison to Confirmatory Factor Analysis and Structural Equation Models, implemented in the `lavaan` `R` package, and Bayesian confirmatory factor analysis, implemented in `blavaan`.


## Simulation Design

For the Bayesian confirmatory factor analysis and structural equation models:

- Did a trial framework with a sparse-loading model (75% of loadings zeroed out) and N  = 100, K = 10, d of 2, 4, 6, 8. The simulation ran for 50 iterations of each pairing.  IRTM showed lower Theta MSE and significantly fewer failed models.

_Limitations on the simulations_
: 
We do not simulate models with a complexity larger than K > 20, d > 8.
IRT-M handles these models relatively straightforwardly, they are extremely computationally intensive and slow in BSEM and BCFA's STAN implementation.  

We show 1-2 cycles of K > 10 to provide an approximate benchmark on time. Recognizing that hours (or weeks) of runtime is impractical for most users, won't do a full 50 samples of each benchmark for the complex models.
The practical upper bound on K is somewhere less than 50 (at K = 250 features, IRTM ran in xx seconds, while BSEM did not start the sampling after two hours of prep. For K = 50 features, BSEM had sampled only one chain after 2 hours.)



## Recovering lambda

Loss Function

We compare model performance in recovering the true lambda loadings via mean squared error (MSE). 

For IRT-M, PCA, and unconstrained IRT models, we standardize the MSE by dividing our predictions by the estimate standard deviations.

For the IRT-M versus CFA and SEM comparisons, we used an unstandardized MSE (because of how CFA and SEM handle non-loading factor-lambda pairs).

```{r pressure, echo=FALSE, eval=FALSE}
mse <- function(ypred, ytrue, aggregate=TRUE, root=FALSE){
  if(length(dim(ypred)) == 3)
    ypred <- apply(ypred, c(1,2), mean)

  ytrue_sd <- apply(ytrue, 2, sd)
  ypred_sd <- apply(ypred, 2, sd)
  if (aggregate)
    ret <- mean(colMeans((ytrue/ytrue_sd-ypred/ypred_sd)^2))
  else
    ret <- (ytrue/ytrue_sd-ypred/ypred_sd)^2
  if (root)
    ret <- sqrt(ret)

  return(ret)
}

```

